### 2016-01-12 model/overflowmcmc.tex
Mostly finished writing up model + sketch of algorithm

\section{The Overflow Model}\label{sec:model}
The overflow model assumes we observe several pieces of data, 1) a noisy measurement of the arrival time of block $k$, $t_k$, measured in minutes since the zero'th block, 2) the number of transactions in block $k$, $M_k - M_{k-1}$ so that $M_k$ is the number of transactions on the entire block chain up to and including block $k$, and 3) the amount of data the transactions require to be stored in the block, $D_k$, measured in MBs, all observed for $k=1,2,\dots,n_{b}$.

\subsection{Measurement error for block arrival times}
Let $m_k = \mathrm{median}(t_{k - 1:11})$ where $t_{k-1:11} = (t_{k-11}, t_{k-10},\dots,t_{k-1})'$. Then by design $t_k > m_k$ for all $k$. Let $\tau_k$ denote the actual arrival time of the $k$'th block. Then for $k=1,2,\dots,n_{b}$ $t_k$ has a truncated normal distribution:
\begin{align*}
\log t_k |t_{1:(k-1)}, \tau_{1:n_{b}},\sigma^2 \sim N(\log\tau_k, \sigma^2)1(t_k > m_k).
\end{align*}

@@

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transaction in the entire chain up to time $\tau$, and let $\lambda(\tau)$ denote some intensity function. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\phi\Psi(\tau)$ and rate parameter $\phi$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\phi[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)], \phi)
\end{align*}
with mean $\Psi(\tau_i + \delta_i) - \Psi(\tau_i)$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\phi\psi_k, \phi).
\end{align*}

In order for the gamma process to be well defined we require $\phi>0$ and $\Psi(\tau)$ to be nondecreasing and right continuous with $\Psi(0)=0$. An easy way to achieve this is to define $\Psi(\tau) = \int_0^\tau\psi(t)dt$ where $\psi(t)$ is continuous and nonnegative. Then $\psi(t)$ is the intensity of the Gamma process, but is also intimately related to the intensity of the Poission process. The mean of $N(\tau + \delta) - N(\tau)$ is 
\begin{align*}
E[N(\tau + \delta) - N(\tau)] &= E\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} = E[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
 &= \Psi(\tau + \delta) - \Psi(\delta) = \int_{\tau}^{\tau + \delta}\psi(t)dt.
\end{align*}
In this way, $\psi(t)$ can be seen as the expected intensity function of the $N(\tau)$ process. The Gamma-Poisson structure also allows for a more flexible model on $N(\tau)$ by allowing for overdispersion. In a Poisson model the mean and variance are the same, but in this case they are proportional:
\begin{align*}
V[N(\tau + \delta) - N(\tau)] &= E\{V[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} + V\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\}\\
&= E[\Lambda(\tau + \delta) - \Lambda(\tau)] + Var[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
&= [\Psi(\tau + \delta) - \Psi(\tau)](1 + 1/\phi) = \frac{\phi + 1}{\phi}\int_{\tau}^{\tau + \delta}\psi(t)dt.
\end{align*}
This allows for overdispersion but not underdispersion since $(\phi + 1)/\phi\in[1,\infty)$ for $\phi\ge 0$.

To complete this portion of the model we need to specify $\psi(t)$. A flexible structure that captures many modeling choices is $\psi(t) = \exp(\bm{z}(t)'\bm{\beta})$ where $\bm{z}(t)' = (z_1(t),\dots,z_p(t))$ is a vector of time-varying covariates continuously differentiable in time and $\bm{\beta}'=(\beta_1,\dots,\beta_p)$ is a vector of regression coefficients, for example $\psi(t) = \exp(\beta_1 + t\beta_2)$ in order to capture a simple trend. In that case we have
\begin{align*}
\Psi(\tau+\delta) - \Psi(\tau) &= \int_{\tau}^{\tau + \delta} e^{\bm{z}(t)'\bm{\beta}}dt = \int_\tau^{\tau + \delta} e^{\beta_1 + t\beta_2}dt = \frac{e^{\beta_1 + \tau\beta_2}}{\beta_2}\left(e^{\delta\beta_2} - 1\right)
\intertext{and thus}
\psi_k &= \frac{e^{\beta_1 + \tau_{k-1}\beta_2}}{\beta_2}\left(e^{\delta_k\beta_2} - 1\right).
\end{align*}
\subsection{Transaction data}
Let $i=1,2,\dots,n_{trans}$ index transactions and $d_i$ denote the amount of data for transaction $i$ measured in MBs. Then we assume that for $i=1,2,\dots,n_{trans}$
\begin{align*}
d_i \stackrel{iid}{\sim} G(\alpha_d,\beta_d)
\end{align*}
with mean $\alpha_d/\beta_d$. Along with the Poisson process for $N(\tau)$, this implies that $D(\tau)$, the amount of transaction data that has arrived by time $\tau$, is a compound non-homogenous Poisson process conditional on $\Lambda(\tau)$.

\subsection{Overflow and observations of $M_k$ and $D_k$}
Each block has an upper limit for the amount of transaction data it can hold defined by the block miner. Denote this upper limit by $\bar{D}_k$ --- because different miners use different limits and the same miner uses different limits at different times, each block has its own upper limit. Then the number of transactions in blocks $1,2,\dots,k$ combined, $M_k$, could potentially be less than the number of transactions that have arrive by the time the $k$'th block arrives, $N_k$. Specifically:
\begin{align*}
M_k &= \max\left\{m\in\left\{M_{k-1}, M_{k-1} + 1, \dots, N_k\right\}: \sum_{i = M_{k-1} + 1}^m d_i \le \bar{D}_k\right\}
\end{align*}
so that
\begin{align*}
D_k &= \sum_{i = M_{k-1} + 1}^{M_k} d_i
\end{align*}
with the convention that empty summations are defined as zero, i.e. $\sum_{i=5}^4d_i \equiv 0$. Finally, the upper limits are modeled as a categorical distribution over a set of prespecified possible limits, $\Delta_1,\dots,\Delta_{n_{limits}}$ with probability $p_{ik} = P(\bar{D}_k = \Delta_i)$, $\sum_ip_{ik} = 1$ and $p_{ik} \ge 0$. In other words
\begin{align*}
\bar{D}_k\stackrel{ind}{\sim} Cat(\Delta_{1:n_{limits}}, p_{1:n_{limits},k}).
\end{align*}

@@

In order to simulate from the posterior we construct a Gibbs sampler using data augmentation to expand the state space and Metropolis steps where appropriate. We use data augmentation in two ways. First, in order to more effectively deal with the truncated distribution in the measurement error model for block arrival times we introduce $\tilde{\bm{t}}_k=(\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)$ as the original $t_k$ along with $\tilde{n}_k$ draws from the untruncated lognormal distribution for $t_k$ that occurred before $t_k$ was drawn from the restricted support. Second, the models for $\eta_{1:n_{b}}$ and $d_{1:n_{trans}}$ induce a model for $(M_{1:n_{b}},D_{1:n_{trans}})$ since the latter are a detrministic function of the former, but this model is not easy work with. Instead, we explicitly use $\eta_{1:n_{b}}$ and $d_{1:n_{trans}}$ in the Gibbs sampler as additional data augmentation.

\subsection{Data augmentation for truncated distributions}
For block $k=1,2,n_{b}$ the measurement error model on block arrival times is
\begin{align*}
\log t_k \sim N(\log\tau_k,\sigma^2)1(t_k > m_k)
\end{align*}
where $m_k = \mathrm{median}(t_{k - 1:11})$ with density 
\begin{align*}
p(\log\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2) \propto \sigma^{-n_{b}}\frac{\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}(\log t_K - \log \tau_k)^2\right]}{\prod_{k=1}^{n_{b}}\left[1 - \Phi\left(\frac{\log m_k - \log\tau_k}{\sigma}\right)\right]}
\end{align*}
where $\Phi(.)$ is the standard normal cdf. Drawing from the full conditional distributions of $\sigma^2$ and $\tau_k$ can be challenging since both parameters enter into the normalizing constant of $p(\log\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2)$. Consider an underlying process which gives rise to the truncated distribution --- instead of simply drawing $t_k$ from the truncated lognormal distribution, this process draws from the untruncated lognormal distribution until it obtains a draw from the restricted region of the parameter space, i.e. until it successfully obtains a $t_k$ such that $t_k > m_k$. Let $\tilde{n}_k$ be the number of failures before the observed success $t_k$. Then $\tilde{n}_k$ has a negative binomial distribution, specifically $\tilde{n}_k \sim NB(1, p_k)$ with density
\begin{align*}
p(\tilde{n}_k|\tau_k,\sigma^2) = p_k^{\tilde{n}_k}(1 - p_k)
\end{align*}
where $p_k$ is the probability of failure, i.e. $p_k = \Phi([\log m_k - \log \tau_k]/\sigma)$. Conditional on $\tilde{n}_k$ the density of the $\tilde{n}_k$ failures, $\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}$, is again truncated lognormal, but this time truncated to the opposite end of the parameter space. That is for $i = 1,\dots,\tilde{n}_k$
\begin{align*}
\log\tilde{t}_{ik}\stackrel{iid}{\sim}N(\log\tau_k,\sigma^2)1(t_k < m_k).
\end{align*}

So the data augmentation step of the Gibbs sampler consists of the following substeps: for $k=1,2,\dots,n_{b}$,
\begin{enumerate}
\item Draw $\tilde{n}_k \sim NB(1,p_k)$ where $p_k = \Phi([\log m_k - \log \tau_k]/\sigma)$.
\item If $\tilde{n}_k > 0$, for $i=1,\dots,\tilde{n}_k$ draw $\tilde{t}_{ik} \sim N(\log\tau_k,\sigma^2)1(t_k < m_k)$ and form $\tilde{\bm{t}}_k = (\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)'$.
\end{enumerate}
For simplicity in the other Gibbs steps define $\tilde{t}_{\tilde{n}_k + 1,k}\equiv t_k$.

It is tempting to imbue $\tilde{n}_k$ and $\tilde{t_{ik}}$ with some interpretation relative to the model, e.g. as is [CITE GELMAN PAPER]. For example, $\tilde{n}_k$ may represent the number of blocks that were rejected by the Bitcoin protocol for having a timestamp earlier than the median of the previous elevent blocks. While tempting this interpretation is not strictly speak correct. When a block is rejected by the protocol a new block does not appear instantaneously since it takes time for another miner to discover the block. So the arrival times and thus measurement error distributions of the rejected block and the accepted block are not the same. While the interpretation is not correct, it is good to be inspired by computational tricks to create better model [CITE SAME GELMAN PAPER] and perhaps something similar to this negative binomial structure can be used to add rejected blocks as a component of the model.

\subsection{measurement error variance}
The full conditional distribution of the measurement error variance, $\sigma^2$, depends only on $\tilde{\bm{t}}_{1:n_{b}}$ and $\tau_{1:n_{b}}$. It is
\begin{align*}
p(\sigma^2|...) \propto \sigma^{-(n_{b} + \sum_{k=1}^{n_{b}}\tilde{n}_k)}\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\log \tilde{t}_{ik} - \log\tau_k)^2\right]p(\sigma^2)
\end{align*}
where $p(\sigma^2)$ is the density of the prior distribution on $\sigma^2$. We will suppose that $\sigma$ has a half-$t$ prior, $\sigma \sim t_\nu(S)1(\sigma>0)$ with degrees of freedom and scale parameter $S$. This prior is equivalent to the marginal prior on $\sigma$ when $\sigma^2|\omega \sim IG(\nu/2, \omega)$ and $\omega \sim G(1/2, 1/(\nu S^2))$. Then letting $a_\sigma = (n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k)/2$ and $b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\log \tilde{t}_{ik} - \log\tau_k)^2$ the full conditional distribution of $(\sigma^2,\omega)$ is
\begin{align*}
p(\sigma^2,\omega|...) \propto (\sigma^2)^{-(a_\sigma+\nu)/2-1}\left[-\frac{b_\sigma}{\sigma^2}\right]\exp\left[-\frac{\omega}{\sigma^2}\right]\omega^{(\nu + 1)/2 - 1}\exp\left[-\frac{\omega}{\nu S^2}\right].
\end{align*}
So we draw $\sigma^2$ and $\omega$ in two separate Gibbs steps:
\begin{enumerate}
\item Draw $\omega \sim G\left(\frac{\nu + 1}{2}, \frac{1}{\sigma^2} + \frac{1}{\nu S^2}\right)$ and form
\begin{align*}
a_\sigma = n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k &&\mbox{  and  }&& b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\log \tilde{t}_{ik} - \log\tau_k)^2/2;
\end{align*}
\item Draw $\sigma^2 \sim  IG([a_\sigma + \nu]/2, b_\sigma + \omega)$.
\end{enumerate}

>>>>

\section{The Overflow Model}\label{sec:model}
The overflow model assumes we observe several pieces of data: 1) a noisy measurement of the arrival time of block $k$ measured in minutes since the zero'th block and denoted by $t_k$, 2) the number of transactions in block $k$ denoted by $M_k - M_{k-1}$ so that $M_k$ is the number of transactions on the entire block chain up to and including block $k$, and 3) the amount of transaction data stored in block $k$ measured in MBs and denoted by $D_k$, all observed for $k=1,2,\dots,n_{b}$.

\subsection{Measurement error for block arrival times}
Let $m_k = \mathrm{median}(t_{k - 1:11})$ where $t_{k-1:11} = (t_{k-11}, t_{k-10},\dots,t_{k-1})'$. Then by design $t_k > m_k$ for all $k$. Let $\tau_k$ denote the actual arrival time of the $k$'th block. Then for $k=1,2,\dots,n_{b}$, $t_k$ has a truncated normal distribution:
\begin{align*}
t_k |t_{1:(k-1)}, \tau_{1:n_{b}},\sigma^2 \sim N(\tau_k, \sigma^2)1(t_k > m_k).
\end{align*}

@@

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transactions that have arrive by time $\tau$, and let $\lambda(\tau)$ denote some intensity function such that $\lambda(\tau) > 0$ for all times $\tau$. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ where $\delta_i>0$ for all $i$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\Psi(\tau)$ and rate parameter $\phi$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\Psi(\tau_i + \delta_i) - \Psi(\tau_i), \phi)
\end{align*}
with mean $[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)]/\phi$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\psi_k, \phi).
\end{align*}

In order for the gamma process to be well defined we require $\phi>0$ and $\Psi(\tau)$ to be nondecreasing and right continuous with $\Psi(0)=0$. An easy way to achieve this is to define $\Psi(\tau) = \int_0^\tau\psi(t)dt$ where $\psi(t)$ is continuous and nonnegative. While $\psi(t)$ can be thought of as the intensity of the Gamma process it is also intimately related to the intensity of the Poission process. The mean of $N(\tau + \delta) - N(\tau)$ is 
\begin{align*}
E[N(\tau + \delta) - N(\tau)] &= E\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} = E[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
 &= [\Psi(\tau + \delta) - \Psi(\delta)]/\phi = \int_{\tau}^{\tau + \delta}\psi(t)dt/\phi.
\end{align*}
In this way, $\psi(t)/\phi$ can be seen as the expected intensity function of the $N(\tau)$ process. The Gamma-Poisson structure also allows for a more flexible model on $N(\tau)$ by allowing for overdispersion. In a Poisson model the mean and variance are the same, but in this case they are proportional:
\begin{align*}
V[N(\tau + \delta) - N(\tau)] &= E\{V[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} + V\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\}\\
&= E[\Lambda(\tau + \delta) - \Lambda(\tau)] + V[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
&= [\Psi(\tau + \delta) - \Psi(\tau)](1/\phi + 1/\phi^2) = \frac{\phi + 1}{\phi}E[N(\tau + \delta) - N(\tau)].
\end{align*}
This allows for overdispersion but not underdispersion since $(\phi + 1)/\phi\in[1,\infty)$ for $\phi\ge 0$.

To complete this portion of the model we need to specify $\psi(t)$. A flexible structure that captures many modeling choices is $\psi(t) = \exp(\bm{z}(t)'\bm{\beta})$ where $\bm{z}(t)' = (z_1(t),\dots,z_p(t))$ is a vector of time-varying covariates continuously differentiable in time and $\bm{\beta}'=(\beta_1,\dots,\beta_p)$ is a vector of regression coefficients, for example $\psi(t) = \exp(\beta_1 + t\beta_2)$ in order to capture a simple trend. In that case we have
\begin{align*}
\Psi(\tau+\delta) - \Psi(\tau) &= \int_{\tau}^{\tau + \delta} e^{\bm{z}(t)'\bm{\beta}}dt = \int_\tau^{\tau + \delta} e^{\beta_1 + t\beta_2}dt = \frac{e^{\beta_1 + \tau\beta_2}}{\beta_2}\left(e^{\delta\beta_2} - 1\right)
\intertext{and thus}
\psi_k &= \frac{e^{\beta_1 + \tau_{k-1}\beta_2}}{\beta_2}\left(e^{\delta_k\beta_2} - 1\right).
\end{align*}
\subsection{Transaction data}
Let $i=1,2,\dots,n_d \equiv N_{n_b}$ index transactions and $d_i$ denote the amount of data for transaction $i$ measured in MBs. Then we assume that for $i=1,2,\dots,n_d$
\begin{align*}
d_i \stackrel{iid}{\sim} G(\alpha_d,\beta_d)
\end{align*}
with mean $\alpha_d/\beta_d$. Along with the Poisson process for $N(\tau)$ this implies that the amount of transaction data that has arrived by time $\tau$, $D(\tau)$, is a compound non-homogenous Poisson process conditional on $\Lambda(\tau)$.

\subsection{Overflow and observations of $M_k$ and $D_k$}
Each block has an upper limit for the amount of transaction data it can hold defined by the block miner. Denote this upper limit by $\bar{D}_k$ --- because different miners use different limits and the same miner uses different limits at different times, each block has its own upper limit. Then the number of transactions in blocks $1,2,\dots,k$ combined, $M_k$, could potentially be less than the number of transactions that have arrived by the time the $k$'th block arrives, $N_k$. Specifically:
\begin{align*}
M_k &= \max\left\{m\in\left\{M_{k-1}, M_{k-1} + 1, \dots, N_k\right\}: \sum_{i = M_{k-1} + 1}^m d_i \le \bar{D}_k\right\}
\end{align*}
so that
\begin{align*}
D_k &= \sum_{i = M_{k-1} + 1}^{M_k} d_i
\end{align*}
with the convention that empty summations are defined as zero, i.e. $\sum_{i=5}^4d_i \equiv 0$.

\section{Priors}
The model has the following unknown parameters: $\sigma^2$, $\phi$, $\beta$, $\alpha_d$, $\beta_d$, and $\bar{D}_k$ for $k=1,2,\dots,n_b$. We will suppose that these parameters are mutually independent in the prior, i.e.
\begin{align*}
p(\sigma^2,\phi,\beta,\alpha_d,\beta_d,\bar{D}_1,\dots,\bar{D}_{n_{b}}) = p(\sigma^2)p(\phi)p(\beta)p(\alpha_d)p(\beta_d)p(\bar{D}_1)\hdots p(\bar{D}_{n_{b}}).
\end{align*}

\subsection{Prior on $\sigma^2$}
The expected size of the measurement error on $t_k$ is controlled by $\sigma^2$ --- the larger $\sigma^2$, the farther away we expect $t_k$ to be from $\tau_k$. We will use a half-$t$ prior on $\sigma$ which can be written as the following prior on $\sigma^2$:
\begin{align*}
\sigma^2|\omega &\sim IG(v/2, \omega); && \omega \sim G(1/2, s^2/2).
\end{align*}
This yields a $t_{v}(0,s^2)1(\sigma>0)$ prior on $\sigma$ where $s^2$ is a scale parameter and $v$ is a degrees of freedom parameter. We use $v=1$ so that the prior on $\sigma$ has a half-Cauchy distribution with median $s$. To gain some intuition in order to set an appropriate value for $s$, suppose that the measurement error distribution was not truncated. Then we would expect $68\%$ of observed block arrival times to be within $\sigma$ of the actual arrival time. A reasonable a priori guess for $\sigma$ is then one minute, so we set $s=1$ so that the median value of $\sigma$ in its prior is one. This prior is only weakly informative --- the right tail is gently sloped so that the data will quickly overwhelm the prior as observations pile up [CITE GELMAN VARIANCE PAPER].

\subsection{Prior on $\phi$}
The conditionally conjugate prior for $\phi$ is a gamma distribution, so we will suppose that $\phi \sim G(a_{\phi}, b_{\phi})$. To choose $a_\phi$ and $b_\phi$ recall that $\phi$ is a dispersion parameter:
\begin{align*}
V[N(\tau)|\phi,\Psi(\tau)] = \frac{\phi + 1}{\phi}E[N(\tau)|\phi,\Psi(\tau)] = \frac{\phi + 1}{\phi}\frac{\Psi(\tau)}{\phi}.
\end{align*}
Setting $\phi = 1$ means that the variance of $N(\tau)$ is twice its mean. Absent strong prior information about $\phi$, we will center our prior on $2$ with a high level of uncertainty. The mean of the gamma distribution is $a_\phi/b_\phi$ with variance $a_\phi/b_\phi^2$, so we will set $b_\phi = 1/100$ to reflect a high degree of uncertainty, and $a_\phi = 1/50$ so that $a_\phi/b_\phi = 2$.

\subsection{Prior on $\bm{\beta}$}
As we will see in the next section, there is no convenient conditionally conjugate form for $\bm{\beta}$, so there is no computational trade-off associated with choosing a prior which more accurately reflects prior knowledge. We will consider $\bm{\beta}$ much like a regression parameter and set
\begin{align*}
\bm{\beta} \sim N(\bm{b}_\beta, \bm{S}_\beta)
\end{align*}
where $\bm{b}_\beta$ is a mean vector and $\bm{S}_\beta$ is a covariance matrix. We will use a weakly informative prior with $\bm{b}_\beta = \bm{0}$ and $\bm{S}_\beta = s^2_\beta\bm{I}$ where $s_\beta = 100$.  [MAYBE SOMETHING WITH FATTER TAILS HERE? MAYBE NOTE THAT NONINFORMATIVE PRIOR IS NOT A GOOD IDEA SINCE WE WANT TO GUARANTEE POSTERIOR PROPRIETY --- SAME FOR $s_\beta$ TOO LARGE (VIVEK'S PAPER)]

\subsection{Priors on $\alpha_d$ and $\beta_d$}

\subsection{Priors on $\bar{D}_k$'s}

Finally, the upper limits are modeled as a categorical distribution over a set of prespecified possible limits, $\Delta_1,\dots,\Delta_{n_{lim}}$ with probability $p_{ik} = P(\bar{D}_k = \Delta_i)$, $\sum_ip_{ik} = 1$ and $p_{ik} \ge 0$. In other words
\begin{align*}
\bar{D}_k\stackrel{ind}{\sim} Cat(\Delta_{1:n_{lim}}, p_{1:n_{lim},k}).
\end{align*}

@@

In order to simulate from the posterior we construct a Gibbs sampler using data augmentation to expand the state space and Metropolis steps where appropriate. We use data augmentation in two ways. First, in order to more effectively deal with the truncated distribution in the measurement error model for block arrival times we introduce $\tilde{\bm{t}}_k=(\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)$ as the original $t_k$ along with $\tilde{n}_k$ draws from the untruncated normal distribution for $t_k$ that occurred before $t_k$ was drawn from the restricted support. Second, the models for $\eta_{1:n_{b}}$ and $d_{1:n_d}$ induce a model for $(M_{1:n_{b}},D_{1:n_d})$ since the latter are a detrministic function of the former, but this model is not easy work with. Instead, we explicitly use $\eta_{1:n_{b}}$ and $d_{1:n_d}$ in the Gibbs sampler as additional data augmentation.

\subsection{Data augmentation for truncated distributions --- $\tilde{n}_k$ and $\tilde{t}_k$}
For block $k=1,2,n_{b}$ the measurement error model on block arrival times is
\begin{align*}
t_k|t_{1:(k-1)} \sim N(\tau_k,\sigma^2)1(t_k > m_k)
\end{align*}
where $m_k = \mathrm{median}(t_{k - 1:11})$ with density 
\begin{align*}
p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2) \propto \sigma^{-n_{b}}\frac{\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}(t_k - \tau_k)^2\right]}{\prod_{k=1}^{n_{b}}\left[1 - \Phi\left(\frac{m_k - \tau_k}{\sigma}\right)\right]}
\end{align*}
where $\Phi(.)$ is the standard normal cdf. Drawing from the full conditional distributions of $\sigma^2$ and $\tau_k$ can be challenging since both parameters enter into the normalizing constant of $p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2)$. Consider an underlying process which gives rise to the truncated distribution --- instead of simply drawing $t_k$ from the truncated normal distribution, this process draws from the untruncated normal distribution until it obtains a draw from the restricted region of the parameter space, i.e. until it successfully obtains a $t_k > m_k$. Let $\tilde{n}_k$ be the number of failures before the observed success $t_k$. Then $\tilde{n}_k$ has a negative binomial distribution, specifically $\tilde{n}_k \sim NB(1, \rho_k)$ with density
\begin{align*}
p(\tilde{n}_k|\tau_k,\sigma^2) = \rho_k^{\tilde{n}_k}(1 - \rho_k)
\end{align*}
where $\rho_k$ is the probability of failure, i.e. $\rho_k = \Phi([m_k - \tau_k]/\sigma)$. Conditional on $\tilde{n}_k$ the density of the $\tilde{n}_k$ failures, $\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}$, is again truncated normal, but this time truncated to the opposite end of the parameter space. That is for $i = 1,\dots,\tilde{n}_k$
\begin{align*}
\tilde{t}_{ik}\stackrel{iid}{\sim}N(\tau_k,\sigma^2)1(t_k < m_k).
\end{align*}
So the data augmentation step of the Gibbs sampler consists of the following substeps: for $k=1,2,\dots,n_{b}$,
\begin{enumerate}
\item Draw $\tilde{n}_k \sim NB(1,\rho_k)$ where $\rho_k = \Phi([ m_k -  \tau_k]/\sigma)$.
\item If $\tilde{n}_k > 0$, for $i=1,\dots,\tilde{n}_k$ draw $\tilde{t}_{ik} \sim N(\tau_k,\sigma^2)1(t_k < m_k)$ and form $\tilde{\bm{t}}_k = (\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)'$.
\end{enumerate}
For simplicity in the other Gibbs steps define $\tilde{t}_{\tilde{n}_k + 1,k}\equiv t_k$.

It is tempting to imbue $\tilde{n}_k$ and $\tilde{t_{ik}}$ with some interpretation relative to the model, e.g. as in [CITE GELMAN PAPER]. For example, $\tilde{n}_k$ may represent the number of blocks that were rejected by the Bitcoin protocol for having a timestamp earlier than the median of the previous elevent blocks. While tempting this interpretation is not strictly speaking correct. When a block is rejected by the protocol a new block does not appear instantaneously since it takes time for another miner to discover the block. So the arrival times and thus measurement error distributions of the rejected block and the accepted block are not the same. While the interpretation is not correct, it is good to be inspired by computational tricks to create better model [CITE SAME GELMAN PAPER] and perhaps something similar to this negative binomial structure can be used to add rejected blocks as a component of the model, though we do not explore this possibility here.

\subsection{Measurement error variance --- $\sigma^2$}
The full conditional distribution of the measurement error variance, $\sigma^2$, depends only on $\tilde{\bm{t}}_{1:n_{b}}$ and $\tau_{1:n_{b}}$. It is
\begin{align*}
p(\sigma^2|...) \propto \sigma^{-(n_{b} + \sum_{k=1}^{n_{b}}\tilde{n}_k)}\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2\right]p(\sigma^2)
\end{align*}
where $p(\sigma^2)$ is the density of the prior distribution on $\sigma^2$. With the half-$t$ prior on $\sigma$ discussed above and letting $a_\sigma = (n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k)/2$ and $b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2$ the full conditional distribution of $(\sigma^2,\omega)$ is
\begin{align*}
p(\sigma^2,\omega|...) \propto (\sigma^2)^{-(a_\sigma+v)/2-1}\exp\left[-\frac{b_\sigma}{\sigma^2}\right]\exp\left[-\frac{\omega}{\sigma^2}\right]\omega^{(v + 1)/2 - 1}\exp\left[-\frac{\omega}{v s^2}\right].
\end{align*}
So we draw $\sigma^2$ and $\omega$ in two separate Gibbs steps:
\begin{enumerate}
\item Draw $\omega \sim G\left(\frac{v + 1}{2}, \frac{1}{\sigma^2} + \frac{1}{v s^2}\right)$ and form
\begin{align*}
a_\sigma = n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k &&\mbox{  and  }&& b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2/2;
\end{align*}
\item Draw $\sigma^2 \sim  IG\left(\frac{a_\sigma + v}{2}, b_\sigma + \omega\right)$.
\end{enumerate}

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\frac{\phi^{\Psi(\tau_k) - \Psi(\tau_{k-1})}}{\Gamma\left[\Psi(\tau_k) - \Psi(\tau_{k-1})\right]}\lambda_k^{\Psi(\tau_k)-\Psi(\tau_{k-1}) - 1}
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the Gamma function. As a result, we are limited to several options for this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\end{enumerate}
Options 2 and 3 could both be replaced with a rejection sampling step under some circumstances, but a Metropolis step is likely cheaper computationally [CITE THE LIU PAPER].
